Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-02-27 manuscript/manuscript.Rmd
updated ms with new figures from 1-steps

```{r libraris, include=FALSE}
library(ggplot2)
library(plyr)
library(gridExtra)
library(reshape2)
```

@@

@@

```{r figure_2, dependson="plot-options", fig.cap="Proportional change in species' mean cover caused by a 1% increase in observed precipitation, temperature, or both as predicted by the individual-based IPM and the aggregate-based QBM. Note that the change for POSE due to a precipitation increase predicted by the QBM is almost zero and so does not show up in the figure. Labels on x-axis refer to: 'ppt' = 1% increase in mean precipitation; 'temp' = 1% increase in mean temperature; 'ppt + temp' = 1% increase in mean precipitation and mean temperature.", fig.width=6.5, fig.height=6, cache=FALSE, results='hide'}
climBase <- readRDS("../montana/quadBM/simulations/baselineSimulationQBM.rds")
climPpt <- readRDS("../montana/quadBM/simulations/pptChangeSimulationQBM.rds")
climTemp <- readRDS("../montana/quadBM/simulations/tempChangeSimulationQBM.rds")
climBoth <- readRDS("../montana/quadBM/simulations/temp-pptChangeSimulationQBM.rds")
climBase$sim <- rep("AObserved", nrow(climBase)) 
climPpt$sim <- rep("BPrecipitation", nrow(climBase)) 
climTemp$sim <- rep("CTemperature", nrow(climBase)) 
climBoth$sim <- rep("DBoth", nrow(climBase)) 
climSimsQBM <- rbind(climBase, climPpt, climTemp, climBoth)
qbm_means <- ddply(climSimsQBM, .(species, sim), summarise,
                   avg_cover = mean(cover))
tmpid <- which(qbm_means$sim=="AObserved")
diffs <- list()
for(i in 1:length(tmpid)){
  obs <- qbm_means[tmpid[i],"avg_cover"]
  diffs[[i]] <- (rep(obs,3) - qbm_means[(tmpid[i]+1):(tmpid[i]+3), "avg_cover"])/rep(obs,3)
}
names(diffs) <- unique(qbm_means$species)
diff_df <- melt(as.data.frame(diffs))
diff_df$sim <- rep(c("pptChange", "tempChange", "temppptChange"),4)
diff_df$type <- rep("QBM", nrow(diff_df))

>>>>

```{r libraris, include=FALSE}
library(ggplot2)
library(plyr)
library(gridExtra)
library(reshape2)
library(xtable)
```

@@

Tables
------
```{r table_1,results='asis',message=FALSE, comment=NA, echo=FALSE}
# Make a table for average residual from each model and year effects structure
ipm_onestep <- readRDS("../montana/ipm/simulations/one_step_results/ipm_one-step_forecasts_combined.rds")
resD <- readRDS("../montana/quadBM/simulations/oneStepResids_YearEffects.rds")
resDny <- readRDS("../montana/quadBM/simulations/oneStepResids_NoYearEffects.rds")
resD$simtype <- rep("full", nrow(resD))
resDny$simtype <- rep("noyear", nrow(resDny))
resDall <- rbind(resD, resDny)
noZ <- which(resDall$lagCover!=0)
resDall <- resDall[noZ,]

stats <- ddply(ipm_onestep, .(species, year), summarise,
               mean_resid = mean(resid),
               sd_resid = sd(resid),
               min_resid = min(resid),
               max_resid = max(resid))
stats$model <- rep("IPM", nrow(stats))
colnames(stats)[2] <- "simtype"
stats2 <- ddply(resDall, .(species, simtype), summarise,
               mean_resid = mean(resids),
               sd_resid = sd(resids),
               min_resid = min(resids),
               max_resid = max(resids))
stats2$model <- rep("QBM", nrow(stats2))
stats$simtype <- stats2$simtype
stats <- rbind(stats, stats2)
stats <- stats[,c(ncol(stats),1:(ncol(stats)-1))]
stats$simtype <- rep(c("Full", "Climate"),(nrow(stats)/2))
colnames(stats) <- c("Model", "Species", "Year Structure", "Mean", "SD", "Min", "Max")
print.xtable(xtable(stats, caption = "Table of residual summary statistics."),type="latex", comment=FALSE, include.rownames=FALSE)
```

@@

```{r figure_2, dependson="plot-options", fig.cap="One step forecasts from IPM version.",fig.width=8, fig.height=8, cache=FALSE, results='hide'}
ipm_onestep <- readRDS("../montana/ipm/simulations/one_step_results/ipm_one-step_forecasts_combined.rds")
ipm_onestep$t1 <- ipm_onestep$t1+1900
p1 <- ggplot(subset(ipm_onestep, species=="BOGR"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(t1),y=resid,fill=year), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols)+
  coord_cartesian(ylim = c(-50, 50))+
  ggtitle(expression(italic("B. gracilis")))+
  guides(fill=FALSE)
p2 <- ggplot(subset(ipm_onestep, species=="HECO"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(t1),y=resid,fill=year), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols, labels=c("Year effects", "No year effects"))+
  ggtitle(expression(italic("H. comita")))+
  coord_cartesian(ylim = c(-10, 5))+
  guides(fill=guide_legend(title=""))+
  theme(legend.position=c(0.5,0.2))
p3 <- ggplot(subset(ipm_onestep, species=="PASM"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(t1),y=resid,fill=year), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols)+
  ggtitle(expression(italic("P. smithii")))+
  guides(fill=FALSE)+
  coord_cartesian(ylim = c(-1.5, 1.5))
p4 <- ggplot(subset(ipm_onestep, species=="POSE"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(t1),y=resid,fill=year), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols)+
  ggtitle(expression(italic("Poa secunda")))+
  guides(fill=FALSE)+
  coord_cartesian(ylim = c(-10, 15))
pFinal <- grid.arrange(p1,p2,p3,p4, ncol=2, nrow=2)
pFinal
```

```{r figure_3, dependson="plot-options", fig.cap="One step forecasts from IPM version.",fig.width=8, fig.height=8, cache=FALSE, results='hide'}
ipm_onestep2 <- subset(ipm_onestep, year=="ayear effect")
resDall2 <- subset(resDall, simtype=="full")
ipm_qbm_onestep <- data.frame(year = c(ipm_onestep2$t1, resDall2$year),
                              resids = c(ipm_onestep2$resid, resDall2$resids),
                              model = c(rep("IPM", nrow(ipm_onestep2)),
                                        rep("QBM", nrow(resDall2))),
                              species = c(ipm_onestep2$species, resDall2$species))
g1 <- ggplot(subset(ipm_qbm_onestep, species=="BOGR"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(year),y=resids,fill=model), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols)+
  coord_cartesian(ylim = c(-50, 50))+
  ggtitle(expression(italic("B. gracilis")))+
  guides(fill=FALSE)
g2 <- ggplot(subset(ipm_qbm_onestep, species=="HECO"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(year),y=resids,fill=model), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols)+
  coord_cartesian(ylim = c(-10, 5))+
  ggtitle(expression(italic("H. comata")))+
  guides(fill=guide_legend(title=""))+
  theme(legend.position=c(0.5,0.2))
g3 <- ggplot(subset(ipm_qbm_onestep, species=="PASM"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(year),y=resids,fill=model), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols)+
  ggtitle(expression(italic("P. smithii")))+
  guides(fill=FALSE)+
  coord_cartesian(ylim = c(-1.5, 1.5))
g4 <- ggplot(subset(ipm_qbm_onestep, species=="POSE"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(year),y=resids,fill=model), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols)+
  ggtitle(expression(italic("Poa secunda")))+
  coord_cartesian(ylim = c(-10, 15))+
  guides(fill=FALSE)
pFinal <- grid.arrange(g1,g2,g3,g4, ncol=2, nrow=2)
pFinal
```

```{r figure_4, dependson="plot-options", fig.cap="Proportional change in species' mean cover caused by a 1% increase in observed precipitation, temperature, or both as predicted by the individual-based IPM and the aggregate-based QBM. Note that the change for POSE due to a precipitation increase predicted by the QBM is almost zero and so does not show up in the figure. Labels on x-axis refer to: 'ppt' = 1% increase in mean precipitation; 'temp' = 1% increase in mean temperature; 'ppt + temp' = 1% increase in mean precipitation and mean temperature.", fig.width=6.5, fig.height=6, cache=FALSE, results='hide'}
climBase <- readRDS("../montana/quadBM/simulations/baselineSimulationQBM.rds")
climPpt <- readRDS("../montana/quadBM/simulations/pptChangeSimulationQBM.rds")
climTemp <- readRDS("../montana/quadBM/simulations/tempChangeSimulationQBM.rds")
climBoth <- readRDS("../montana/quadBM/simulations/temp-pptChangeSimulationQBM.rds")
climBase$sim <- rep("AObserved", nrow(climBase)) 
climPpt$sim <- rep("BPrecipitation", nrow(climBase)) 
climTemp$sim <- rep("CTemperature", nrow(climBase)) 
climBoth$sim <- rep("DBoth", nrow(climBase)) 
climSimsQBM <- rbind(climBase, climPpt, climTemp, climBoth)
qbm_means <- ddply(climSimsQBM, .(species, sim), summarise,
                   avg_cover = mean(cover))
tmpid <- which(qbm_means$sim=="AObserved")
diffs <- list()
for(i in 1:length(tmpid)){
  obs <- qbm_means[tmpid[i],"avg_cover"]
  diffs[[i]] <- (rep(obs,3) - qbm_means[(tmpid[i]+1):(tmpid[i]+3), "avg_cover"])/rep(obs,3)
}
names(diffs) <- unique(qbm_means$species)
diff_df <- melt(as.data.frame(diffs))
diff_df$sim <- rep(c("pptChange", "tempChange", "temppptChange"),4)
diff_df$type <- rep("QBM", nrow(diff_df))

### 2015-02-27 manuscript/manuscript.tex
updated ms with new figures from 1-steps

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_2.pdf}
\caption{Proportional change in species' mean cover caused by a 1\%
increase in observed precipitation, temperature, or both as predicted by
the individual-based IPM and the aggregate-based QBM. Note that the
change for POSE due to a precipitation increase predicted by the QBM is
almost zero and so does not show up in the figure. Labels on x-axis
refer to: `ppt' = 1\% increase in mean precipitation; `temp' = 1\%
increase in mean temperature; `ppt + temp' = 1\% increase in mean
precipitation and mean temperature.}
\end{figure}

>>>>

\subsection{Tables}\label{tables}

\begin{table}[ht]
\centering
\begin{tabular}{lllrrrr}
  \hline
Model & Species & Year Structure & Mean & SD & Min & Max \\ 
  \hline
IPM & BOGR & Full & -0.24 & 5.78 & -46.52 & 43.05 \\ 
  IPM & BOGR & Climate & -0.28 & 11.40 & -230.10 & 47.81 \\ 
  IPM & HECO & Full & -0.29 & 0.95 & -10.99 & 4.33 \\ 
  IPM & HECO & Climate & 0.14 & 1.05 & -8.65 & 6.67 \\ 
  IPM & PASM & Full & -0.04 & 0.30 & -4.29 & 1.04 \\ 
  IPM & PASM & Climate & 0.05 & 0.20 & -2.04 & 0.99 \\ 
  IPM & POSE & Full & -0.58 & 1.65 & -30.73 & 7.34 \\ 
  IPM & POSE & Climate & 0.23 & 1.51 & -6.08 & 14.32 \\ 
  QBM & BOGR & Full & -2.06 & 12.45 & -81.48 & 52.88 \\ 
  QBM & BOGR & Climate & -1.86 & 13.61 & -94.71 & 54.33 \\ 
  QBM & HECO & Full & -0.25 & 2.28 & -68.53 & 7.84 \\ 
  QBM & HECO & Climate & -0.47 & 3.49 & -94.57 & 7.91 \\ 
  QBM & PASM & Full & -0.04 & 0.50 & -15.13 & 1.56 \\ 
  QBM & PASM & Climate & -0.05 & 0.53 & -9.69 & 1.67 \\ 
  QBM & POSE & Full & -0.05 & 2.07 & -58.34 & 13.97 \\ 
  QBM & POSE & Climate & -0.01 & 2.47 & -73.14 & 15.18 \\ 
   \hline
\end{tabular}
\caption{Table of residual summary statistics.} 
\end{table}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_2.pdf}
\caption{One step forecasts from IPM version.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_3.pdf}
\caption{One step forecasts from IPM version.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_4.pdf}
\caption{Proportional change in species' mean cover caused by a 1\%
increase in observed precipitation, temperature, or both as predicted by
the individual-based IPM and the aggregate-based QBM. Note that the
change for POSE due to a precipitation increase predicted by the QBM is
almost zero and so does not show up in the figure. Labels on x-axis
refer to: `ppt' = 1\% increase in mean precipitation; `temp' = 1\%
increase in mean temperature; `ppt + temp' = 1\% increase in mean
precipitation and mean temperature.}
\end{figure}
